The genomic impacts of drift and selection for hybrid performance in maize

### 2015-08-19 response.tex

Figure 3 is quite useful and helps to show large physical regions that have lost heterozygosity.  The authors discuss the physical-genetic difference.  It would be useful to have a supplemental figure that shows these same values (cycle 16 heterozygosity on y-axis) versus genetic position on x-axis.  I am curious if there are ever regions of multiple cM on chromosome arms that become fixed.  This would be a nice complementary figure.  These are provided for some chromosomes in 4B and 5B but would be useful to see genome wide.

\response{Heterozygosity plotted on a genetic map was shown in the original supplemental figure 2 (which the reviewer pointed out we failed to reference).  We have switched this figure for the original figure 2. As shown in the new figure 3, there are a few regions on chromosome arms that have been fixed and span multiple cM. \jri{do we need to note in the main text? an interesting observation, maybe implying rapid selection?}}

The large regions of fixation are quite intriguing.  Is there any evidence for major flowering time or disease resistance QTL in these regions?

\response{We agree these regions are of interest, but the large size of these regions is such that there is almost certainly overlap with flowering time or resistance QTL.  We would be hesitant to draw any conclusions from such overlap as similar overlap is likely to occur for a number of other traits of less interest, and evaluating significance or enrichment would be difficult. We have generally avoided GO or other enrichment analyses as the aim of the paper was not in identifying specific candidate genes or characterizing their function.}

@@

Also related to the simulations, there are multiple references in the text to empirical results as "deviation from 
simulation" (e.g., p 15) when I suspect what the authors wish to say is that the results differ from neutral 
expectations based on simulations.

\response{We have removed the phrase ``deviation from simulation'' throughout, replacing it with phrases mentioning differences from neutral simulations.}

The other major weakness of the paper is that relatively little background on the experimental populations 
is provided. There is a long list of citations of prior work on this experimental population, but little 
exposition.

\response{We have mentioned the RRS method and populations a bit more in the introduction and included new paragraphs in the methods explaining the RRS method and populations in increased detail.}

@@

A better presentation of the specific target of selection is needed (I infer that selection is on 
improved yield through increased hybrid performance). Ultimately in the Discussion, there is additional 
discussion of the phenotypes, but the presentation is minimal (again pointing to multiple citations) and 
provides little insight into the nature of the selection observed. Better discussion of phenotypes potentially 
targeted by selection is essential if the manuscript is to end, as it does, with a call for further dissection of 
targets of selection in this population.

\response{We have noted in the new paragraphs in the introduction and methods that yield was the main target of selection but that other traits (moisture content at harvest and reduced root and stalk lodging) were also selected upon.}

@@

\response{This was an oversight on our part; we have now incorporated citation to some of this literature and hopefully clarified the discussion.}

Finally, the Discussion seems a bit overly long and could potentially be more tightly written.

\response{We have endeavored to streamline the discussion where possible while incorporating additional citations (as above) and discussion of load as mentioend by reviewer 1.}

Minor points: 
- Careful copy editing is needed. Heterozygsoity in the figures and figure captions is referred to variously 
as "H" with and without italics and also as heterozygosity spelled out, sometimes within the same figure or 
figure caption. Similarily, Fst needs to be presented in a consistent manner. 
Also edit citations for consistent capitalization and in=text citations for consistent spacing.

\response{We have moved the ms from word to latex; we believe this has fixed many of these issues. We have remade figures and hopefully are now consistent in use of Fst and H. We have also corrected bibtex entries in several places to improve consistency of capitalization.}

### 2015-08-19 main.tex
typos, thank you kate

\abstract{
Modern maize breeding relies upon selection in inbreeding populations to improve the performance of cross-population hybrids. 
The United States Department of Agriculture reciprocal recurrent selection experiment between the Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No. 1 (BSCB1) populations represents one of the longest standing models of selection for hybrid performance. 
To investigate the genomic impact of this selection program, \rev{we genotyped the progenitor lines and over 600 individuals across multiple cycles of selection using a genome-wide panel of $\sim 40,000$ SNPs.
We first confirm previous results showing a steady temporal decrease in genetic diversity within populations and a corresponding increase in differentiation between populations.  
Thanks to detailed historical information on experimental design, we were able to perform extensive simulations using founder haplotypes to replicate the experiment in the absence of selection.
These simulations demonstrate that while most of the observed reduction in genetic diversity can be attributed to genetic drift,   
heterozygosity in each population has fallen more than expected. 
We then take advantage of our high-density genotype data to identify extensive regions of haplotype fixation and trace haplotype ancestry to single founder inbred lines.
The vast majority of regions showing such evidence of selection differe between the two populations, providing evidence against the overdominance model as a means of explaining selection for hbyrid vigor. }
We discuss how this pattern is likely to occur during selection for hybrid performance, and how it poses challenges for dissecting the impacts of modern breeding and selection on the maize genome. }

@@

The observation that the same targets of selection are not observed in the opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids. 
Classic overdominance models of heterosis predict that at a single locus, two distinct alleles confer heterozygote advantage when combined, while the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance). 
In the case of true overdominance, we selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 
Because most deleterious alleles are rare in both heterotic groups \citep{Mezmouk2014}, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well, such that selection will have little impact on diversity in the second population. 
Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general \citep{van2012historical}.

>>>>

\abstract{
Modern maize breeding relies upon selection in inbreeding populations to improve the performance of cross-population hybrids. 
The United States Department of Agriculture reciprocal recurrent selection experiment between the Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No. 1 (BSCB1) populations represents one of the longest standing models of selection for hybrid performance. 
To investigate the genomic impact of this selection program, \rev{we genotyped the progenitor lines and over 600 individuals across multiple cycles of selection using a genome-wide panel of $\sim 40,000$ SNPs.
We first confirm previous results showing a steady temporal decrease in genetic diversity within populations and a corresponding increase in differentiation between populations.  
Thanks to detailed historical information on experimental design, we were able to perform extensive simulations using founder haplotypes to replicate the experiment in the absence of selection.
These simulations demonstrate that while most of the observed reduction in genetic diversity can be attributed to genetic drift,   
heterozygosity in each population has fallen more than expected. 
We then take advantage of our high-density genotype data to identify extensive regions of haplotype fixation and trace haplotype ancestry to single founder inbred lines.
The vast majority of regions showing such evidence of selection differ between the two populations, providing evidence against the overdominance model as a means of explaining selection for hybrid vigor. }
We discuss how this pattern is likely to occur during selection for hybrid performance, and how it poses challenges for dissecting the impacts of modern breeding and selection on the maize genome. }

@@

The observation that the same targets of selection are not observed in the opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids. 
Classic overdominance models of heterosis predict that at a single locus, two distinct alleles confer heterozygote advantage when combined, while the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance). 
In the case of true overdominance,  selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 
Because most deleterious alleles are rare in both heterotic groups \citep{Mezmouk2014}, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well, such that selection will have little impact on diversity in the second population. 
Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general \citep{van2012historical}.
